# Spot-deconvolution of spTotal data using BayesPrism + scMuscle
## David W. McKellar
```{r}
library(Seurat, quietly = T)
library(dplyr, quietly = T)
library(ggplot2, quietly = T)
library(patchwork, quietly = T)

# TED, a.k.a. BayesPrism
## https://github.com/Danko-Lab/TED
library(TED, quietly = T)

source("/home/dwm269/DWM_utils/plotting_utils/scThemes.R")
source("/home/dwm269/DWM_utils/sc_utils/seurat_helpers/seutils.R")
source("/home/dwm269/DWM_utils/sc_utils/seurat_helpers/seuplots.R")
```

## Figure settings, colors, & themes
```{r}
# fonts, sizes, etc.
small.font = 6*2
big.font = 8*2
line.width = 0.5
pt.size=0.01
pt.stroke=0.3
label.size=2

source("/home/dwm269/DWM_utils/plotting_utils/scThemes.R")
scTheme <- scThemes(
  small.font = small.font,
  big.font = big.font,
  line.width = line.width,
  pt.size=pt.size,
  pt.stroke=pt.stroke,
  label.size=label.size
)

print(names(scTheme))

# colors
mckolors <- read.csv("/home/dwm269/DWM_utils/plotting_utils/McKolors_v1.csv") %>% 
  as.list() %>%
  lapply(
    FUN=function(X) X[X!=""]
  )
names(mckolors)
```

# Load and subset preprocessed Visium & spTotal data
```{r}
load("/local/workdir/dwm269/totalRNA/spTotal/robjs/skm_list_v2.RData")
meta_vis <- read.csv("/workdir/dwm269/totalRNA/spTotal/resources/metadata_sheets/meta_sheet_visium.csv")
meta_skm <- meta_vis[meta_vis$tissue=="muscle"&meta_vis$rnase_inhib%in%c("ctrl", "Protector"),]

# skm.list <- vis.list[meta_vis$tissue=="muscle"&meta_vis$rnase_inhib%in%c("ctrl", "Protector")]
meta_skm
```


# Load single-cell reference dataset
```{r}
if(!exists("scMuscle.seurat")){
  load("/local/workdir/dwm269/muscle_data/robjs/scMuscle_mm10_clean_v40.RData")
}
head(scMuscle.seurat)
```

# Function for generating an average expression model for a Seurat object
```{r get_cell_type_model}
get_cell_type_model <-
  function(
    SEU, #seurat object
    assay='RNA',
    slot='data',
    cluster.names, #cell type model; pass in metadata column name for desired cell types
    ignore.clusters=NULL, # vector of chars; cluster IDs to ignore in model generation
    cells.use=NULL, # vector of cell barcodes to use; works in combination with ignore.clusters, doublet.classifications
    genes.ignore=NULL, # genes to remove from model
    doublet.classifications=NULL, # name of doublet classification metadata column; default is to ignore classifications
    # nCores=1, #TODO: parallelize with dofor
    verbose=T
  ){
    require(Seurat)

    #build singlet_cell type_gene expression matrix
    ref.mat <- list() #initialize reference expression matrix
    exp.mat <- GetAssayData(SEU, assay=assay, slot=slot) # Pull out expression data

    # subset based on desired cells
    if(!is.null(cells.use)){
      exp.mat <- exp.mat[,cells.use]
    }

    # remove ignored genes
    if(!is.null(genes.ignore)){
      exp.mat = exp.mat[!rownames(SEU)%in%genes.ignore,]
    }

    #Grab meta.data, filtered for singlets
    if(is.null(cells.use)){ #TODO- subset based on cells.use
      meta <- SEU@meta.data  #ignore sinlget/doublet classifications
    }else{
      meta <- SEU@meta.data[cells.use,]
    }

    cell.types <- sort(unlist(unique(meta[[cluster.names]]))) #Get cell type names
    if(!is.null(ignore.clusters)){
      cell.types <- cell.types[(!cell.types%in%ignore.clusters)]
    }

    # Remove NAs from exp.mat and meta
    exp.mat <- exp.mat[,!is.na(meta[[cluster.names]])]
    meta <- meta[!is.na(meta[[cluster.names]]),]

    # Generate cell type references, one cell type at a time
    if(verbose){cat("Building reference profiles... \n")}
    if(verbose & !is.null(cells.use)){cat("   Using the provided cells... \n")}

    for(i in 1:length(cell.types)){
      tmp.cells = rownames(meta)[ as.vector(meta[[cluster.names]])==cell.types[i] ]

      if(verbose){cat(cell.types[i]," (",i,"/",length(cell.types),")","... ", sep="")}

      if(length(tmp.cells)==0){
        if(verbose){cat(" NO CELLS", sep="")}
      }

      if(length(tmp.cells)==1){
        if(verbose){cat(" Only 1 cell!", sep="")}
        ref.mat[[i]] <- exp.mat[,rownames(meta)%in%tmp.cells]  # grab expression data
      }else{
        if(verbose){cat(length(tmp.cells)," cells.", sep="")}
        ref.mat[[i]] <-
          Matrix::rowMeans( #TODO: should these expression profiles be built on row means?
            exp.mat[,rownames(meta)%in%tmp.cells]  # grab expression data

          )
      }

      if(verbose){cat("\n")}
    }
    ref.mat <- do.call(cbind, ref.mat)
    colnames(ref.mat) <- cell.types

    if(verbose){cat("Done!\n")}

    return(ref.mat)
  }

```

# remove mito and ribo genes; important for deconvolution - 
##     See vignette at https://github.com/Danko-Lab/TED
```{r}
# remove mito and ribo genes
genes.remove <-  unique(c(
  rownames(scMuscle.seurat)[grep("Rps",x = rownames(scMuscle.seurat))], # ribo, short
  rownames(scMuscle.seurat)[grep("Rpl",x = rownames(scMuscle.seurat))], # ribo long 
  rownames(scMuscle.seurat)[grep("mt-",x = rownames(scMuscle.seurat))] # mito
))
print(genes.remove)

#Build cell type gene expression profile - see helper_functions_v1.R for details
celltype.gep <- scMuscle.seurat %>% get_cell_type_model(
  slot="counts",
  cluster.names = "harmony_PHATE_IDs",
  genes.ignore = genes.remove 
)
```

# Deconvolution with differentially expressed 
Load in DEGs for scMuscle, filter genes list, and filter gep
```{r}
all.markers <- read.csv("/workdir/dwm269/muscle_data/scMuscle/supplemental_data/scMuscle_harmonytypes_plus_phatebins_markers_SupFile2.csv")

genes.keep = unique(all.markers$gene[all.markers$pct.1>0.5 & abs(all.markers$avg_logFC)>1])
genes.keep = genes.keep[!genes.keep %in% genes.remove]

celltype.gep <- celltype.gep[rownames(celltype.gep) %in% genes.keep, ]
dim(celltype.gep)

gc()
```

Rename genes that are different between mm10 & mm39 (just Mymk and Mymx)
```{r}
rownames(celltype.gep)[rownames(celltype.gep)=="Tmem8c"] <- "Mymk"
rownames(celltype.gep)[rownames(celltype.gep)=="Gm7325"] <- "Mymx"

print(table(rownames(celltype.gep) %in% Features(skm.list[[1]],assay="STARsolo_collapsed")))
```


Run the deconvolution
```{r}
n.cores=60
celltype.bp <- list()
for(i in 1:length(skm.list)){
  cat("Sample: ", meta_skm$sample[i],"\n")
  celltype.bp[[i]] <-run.Ted( 
    ref.dat = t(celltype.gep), #cell types x genes
    X=t(as.matrix(
      GetAssayData(skm.list[[i]],slot="counts",assay="STARsolo_collapsed")
      )), # spot to deconvolve
    cell.type.labels=colnames(celltype.gep),
    input.type = "GEP", 
    n.cores=n.cores
  )
  # gc()
}
```


```{r}
# Add outputs to Seurat objects ####
for(i in 1:length(skm.list)){ # add theta (composition) values as an Assay
  colnames(celltype.bp[[i]]$res$final.gibbs.theta) <- stringr::str_replace_all(rownames(celltype.bp[[i]]$para$input.phi),pattern = " ", replacement="_")
  skm.list[[i]][["celltype.bp"]] <- CreateAssayObject(data=t(celltype.bp[[i]]$res$final.gibbs.theta))
}
```

## Save skm.list
```{r}
save(
  skm.list,
  file="/workdir/dwm269/totalRNA/spTotal/robjs/skm_vis_BP_scMuscle.RData"
)
```


## Plot predicted cell type distributions
```{r fig.height=10, fig.width=8, message=FALSE, warning=FALSE}
# Plot out BP theta values
tmp.titles = stringr::str_remove_all(meta_skm$sample,pattern = "Vis_") %>%
  stringr::str_remove_all(pattern = "_TA") %>%
  stringr::str_remove_all(pattern = "-SkM")

suppressMessages(
  visListPlot(
    seu.list = skm.list,
    assay = "celltype.bp",
    sample.titles = tmp.titles,
    pt.size = 0.6,
    nrow = 1,
    min.value = 10^-2,
    colormap.same.scale = T,
    colormap = rev(RColorBrewer::brewer.pal(n=11,name="Spectral")),
    # features=GetAssayData(skm.list[[1]],assay="celltype.bp")%>% rownames()%>%unlist()
    features = c(
      "FAPs-(Pro-remodeling)",
      "Monocyte-(Patrolling)",
      "Quiescent-MuSCs",
      "Activated-MuSCs",
      "Committed-Myoblasts",
      "Fusing-Myocytes",
      "Myonuclei-(Type-IIb)",
      "Myonuclei-(Type-IIx)"
    )
  )&coord_fixed(1/1.6)
)
```

# Find genes which correlate at the pseudo-bulk level between scMuscle & the yPAP-Visium samples
```{r}
if(!exists("scMuscle.seurat")){
  load("/local/workdir/dwm269/muscle_data/robjs/scMuscle_mm10_clean_v40.RData")
}

scMuscle.avg <- Seurat::AverageExpression(
  scMuscle.seurat,
  assays = "RNA",
  group.by = "orig.ident"
  # group.by = 'harmony_PHATE_IDs'
) %>%
  as.data.frame()
scMuscle.avg$genes <- rownames(scMuscle.avg)

vis.avg <- Seurat::AverageExpression(
  vis.list[[2]],
  assays = "STARsolo_collapsed",
  group.by = "orig.ident"
  # group.by = 'harmony_PHATE_IDs'
) %>% 
  as.data.frame()
vis.avg$genes <- rownames(vis.avg)

# merge 
avg.df <- merge(
  x = scMuscle.avg, 
  y = vis.avg,
  by="genes"
)

gc()
```

```{r compute_correlation}

# Subset out genes used in BP decon
avg.df.sub <- subset(
  avg.df,
  subset=avg.df$genes %in% genes.keep
)

#Subset according to expression values
avg.df.sub <- subset(
  avg.df.sub,
  subset = avg.df.sub$all.y>0.1
)

cor.mat <- cor(
  x=avg.df.sub$all.x,
  y=avg.df.sub$all.y,
  method="pearson"
)
cor.mat
```

```{r plot_correlation}
ggplot(
  avg.df.sub,
  aes(
    x=all.x,
    y=all.y
  )
)+
  geom_point(
    alpha=0.2,
    size=0.5
  )+
  geom_abline()+
  geom_text(
    aes(label=genes),
    size=2
  )+
  # lims(
  #   x=c(0,2),
  #   y=c(0,2)
  # )+
  # scale_x_log10()+scale_y_log10()+
  scTheme$fig1bcd+
  labs(
    x="scMuscle",
    y="spTotal"
  )

```

# Merge and cluster based on BP theta values
```{r}
strs.merged <- merge(
  x=skm.list[[1]],
  y=skm.list[2:length(skm.list)],
  add.cell.ids = meta_skm$sample
)
```

preprocess using the BayesPrism outputs
```{r}
strs.merged@active.assay <- "celltype.bp"
strs.merged <- strs.merged %>%
  ScaleData() %>%
  RunPCA(
    assay="celltype.bp",
    features=Features(strs.merged,assay="celltype.bp"),
    reduction.key = "BayesPrism.PC_",
    verbose=F,
    npcs=10
  )
```

Plot clusters on PCA
```{r}
lapply(
  c("injury.zones","timepoint","polyA"),
  FUN=function(GROUP) DimPlot(
  strs.merged,
  cells=sample(Cells(strs.merged)),
  reduction = "pca",
  group.by=GROUP
)+
  scTheme$umap
)%>% wrap_plots()
```


# Session Info
```{r}
sessionInfo()
```