
```{r warning=FALSE}
# Load the requisite packages and some additional helper functions.
library(Seurat);
packageVersion("Seurat")
library(Matrix); library(stringr); library(dplyr); library(reticulate)
library(readr); library(fitdistrplus); library(ggplot2)
library(nichenetr)
library(EnhancedVolcano)
library(clusterProfiler)
library(patchwork)
# library(AnnotationHub)
# library(org.Mm.eg.db)
# library(tidyverse)
# library(biomaRt) 

```

```{r}
source("./seutils-v2.R")
# source("./../DWM_utils/sc_utils/seurat_helpers/seutils.R")
source("./../DWM_utils/sc_utils/seurat_helpers/seuplots.R")
```


## Read in metadata
```{r}
meta_heart <- read.csv("./../spTotal/resources/meta_sheet_visium.csv")
meta_heart <- meta_heart[meta_heart$tissue == "heart", ]
```


## Load all gene info for mouse reference
```{r}
## Read in mouse gene info
gene_info_gtf <- read.csv("./../resources/gene_lists/GRCm39_GENCODEm28_gene_info_gtf.tsv", sep = "\t")
# gene_info_gtf$GeneSymbol[duplicated(gene_info_gtf$GeneSymbol)]
# gene_info_gtf$mgi_symbol <- ifelse((gene_info_gtf$mgi_symbol == ""), gene_info_gtf$ensembl_gene_id, gene_info_gtf$mgi_symbol)
# gene_info_gtf <- gene_info_gtf[!duplicated(gene_info_gtf$mgi_symbol), ]
table(gene_info_gtf$Biotype)
gene_info_gtf$biotype_groups = gene_info_gtf$Biotype
gene_info_gtf$biotype_groups[gene_info_gtf$biotype_groups %in% grep(pattern = "pseudogene", x = unique(gene_info_gtf$biotype_groups), value = TRUE)] <- "pseudogene"
gene_info_gtf$biotype_groups[gene_info_gtf$biotype_groups %in% grep(pattern = "IG_", x = unique(gene_info_gtf$biotype_groups), value = TRUE)] <- "IG_genes"
gene_info_gtf$biotype_groups[gene_info_gtf$biotype_groups %in% grep(pattern = "TR_", x = unique(gene_info_gtf$biotype_groups), value = TRUE)] <- "TCR_genes"
table(gene_info_gtf$biotype_groups)
```

## Load in the dataset
```{r fig.height=4.5, fig.width=4.0}
load("/workdir/dwm269/totalRNA/spTotal/robjs/heart_list_v3.RData")
cat("Done")

heart.list <- lapply(
  heart.list,
  FUN = function(SEU) FindClusters(
    SEU, 
    res = 1.0,
    verbose=F
  )
)

```


```{r}

# heart.list <- lapply(
#   heart.list,
#   FUN = function(SEU) seuPreProcess(
#     SEU, 
#     assay="kallisto_collapsed", 
#     nfeatures = 10000,
#     res = 0.6,
#     verbose=F
#   )
# )

```
   
   
   
## Add biotype properties
```{r}

# for (i in 1:length(heart.list)) {
#   print(dim(heart.list[[i]]))
#   for (biotype in unique(gene_info_gtf$biotype_groups)) {
#     genelist = intersect(rownames(heart.list[[i]]), gene_info_gtf$GeneSymbol[gene_info_gtf$biotype_groups == biotype])
#     if (length(genelist) >= 2){
#       print(biotype)
#       heart.list[[i]]@meta.data[, paste0("nCount_", biotype, sep = "")] <-
#         colSums(GetAssayData(heart.list[[i]], assay = "kallisto_collapsed", slot = "counts")[genelist, ])
#       heart.list[[i]]@meta.data[, paste0("Percent_", biotype, sep = "")] <-
#         colSums(GetAssayData(heart.list[[i]], assay = "kallisto_collapsed", slot = "counts")[genelist, ]) /
#         colSums(GetAssayData(heart.list[[i]], assay = "kallisto_collapsed", slot = "counts"))
#       heart.list[[i]]@meta.data[, paste0("nFeature_", biotype, sep = "")] <-
#         colSums(GetAssayData(heart.list[[i]], assay = "kallisto_collapsed", slot = "counts")[genelist, ] > 0)
#     }
#   }
# }

```

```{r}
visListPlot(
  heart.list,#[1:2],
  sample.titles = stringr::str_remove_all(meta_heart$sample,pattern = "Vis_") %>%
    # stringr::str_remove_all(pattern ="yPAP-")%>%
    # stringr::str_remove_all(pattern ="ctrl-")%>%
    stringr::str_remove_all(pattern ="Heart-"),
  assay="STARsolo_collapsed",
  reduction="space",
  # slot = 'counts',
  pt.size=0.8,
  legend.position = "bottom",
  font.size = 12,
  axis.title.angle.y=0,
  ncol = 4,
  # combine = F,
  verbose=F,
  features = c("Percent_miRNA", "Percent_snoRNA", "Percent_snRNA", "Percent_protein_coding")
  # reo.genes
) & coord_fixed(ratio = 1.6)

```

## Plot grid of DimPlots on spatial reduction
```{r}

DimPlotGrid <-
  function(heart.list,
           key = "kallisto_collapsed_snn_res.0.6",
           legend.position = "right") {
    (
      DimPlot(heart.list[[1]], reduction = "space", group.by = key) |
        DimPlot(heart.list[[2]], reduction = "space", group.by = key)
    ) /
      (
        DimPlot(heart.list[[3]], reduction = "space", group.by = key) |
          DimPlot(heart.list[[4]], reduction = "space", group.by = key)
      ) &
      theme_void() &
      theme(legend.position = legend.position) &
      coord_fixed(ratio = 1.8) & ggtitle(NULL)
  }

```


```{r fig.height=6, fig.width=6}
DimPlotGrid(heart.list, key = "kallisto_collapsed_snn_res.1", legend.position = "right")

```


```{r}
Idents(heart.list[[1]]) <-
  heart.list[[1]]$kallisto_collapsed_snn_res.1
heart.list[[1]] <-
  RenameIdents(
    heart.list[[1]],
    "0" = "Ventricle",
    "2" = "Atria",
    "1" = "Ventricle",
    "3" = "Cavity",
    "4" = "Ventricle",
    "5" = "Cavity",
    "6" = "Cavity"
  )
heart.list[[1]]$AnatomicalRegion <- Idents(heart.list[[1]])

Idents(heart.list[[2]]) <-
  heart.list[[2]]$kallisto_collapsed_snn_res.1
heart.list[[2]] <-
  RenameIdents(
    heart.list[[2]],
    "1" = "Border zone",
    "2" = "Inflamed atria",
    "6" = "Inflamed atria",
    "0" = "Inflamed ventricle",
    "3" = "Cavity",
    "4" = "Myocarditic region",
    "5" = "Inflamed ventricle"
  )
heart.list[[2]]$AnatomicalRegion <- Idents(heart.list[[2]])

Idents(heart.list[[3]]) <-
  heart.list[[3]]$kallisto_collapsed_snn_res.1
heart.list[[3]] <-
  RenameIdents(
    heart.list[[3]],
    "0" = "Ventricle",
    "3" = "Atria",
    "1" = "Ventricle",
    "2" = "Cavity",
    "4" = "Cavity"
  )
heart.list[[3]]$AnatomicalRegion <- Idents(heart.list[[3]])

Idents(heart.list[[4]]) <-
  heart.list[[4]]$kallisto_collapsed_snn_res.1
heart.list[[4]] <-
  RenameIdents(
    heart.list[[4]],
    "3" = "Inflamed ventricle",
    "0" = "Border zone",
    "1" = "Myocarditic region",
    "4" = "Inflamed atria",
    "5" = "Inflamed atria",
    "6" = "Cavity",
    "7" = "Cavity",
    "8" = "Cavity",
    "2" = "Inflamed ventricle",
    "9" = "Inflamed ventricle"
  )
heart.list[[4]]$AnatomicalRegion <- Idents(heart.list[[4]])
cat("Done")

```

## Create a merged object and fix levels for anatomical region 
```{r}

heart.list[[1]]$Sample = "Control-Mock"
heart.list[[2]]$Sample = "Control-Infected"
heart.list[[3]]$Sample = "yPAP-Mock"
heart.list[[4]]$Sample = "yPAP-Infected"

heart.list.merged = merge(x = heart.list[[1]],
                          y = heart.list[c(2:4)],
                          add.cell.ids = meta_heart$sample)

heart.list.merged$AnatomicalRegion = factor(
  heart.list.merged$AnatomicalRegion,
  levels = c(
    "Ventricle",
    "Atria",
    "Inflamed ventricle",
    "Inflamed atria",
    "Myocarditic region",
    "Border zone",
    "Cavity"
  )
)

heart.list[[1]]$AnatomicalRegion = factor(heart.list[[1]]$AnatomicalRegion,
                                          levels = levels(heart.list.merged$AnatomicalRegion))
heart.list[[2]]$AnatomicalRegion = factor(heart.list[[2]]$AnatomicalRegion,
                                          levels = levels(heart.list.merged$AnatomicalRegion))
heart.list[[3]]$AnatomicalRegion = factor(heart.list[[3]]$AnatomicalRegion,
                                          levels = levels(heart.list.merged$AnatomicalRegion))
heart.list[[4]]$AnatomicalRegion = factor(heart.list[[4]]$AnatomicalRegion,
                                          levels = levels(heart.list.merged$AnatomicalRegion))

cat("Done")
```  


```{r fig.height=5, fig.width=6}

cp <- c("#5AA878", "#E9DF4F", "#3670B7", "#D85B2A", "#C66FA9", "#BDBEC0", "#E3932E", "#60B8EF")
DimPlotGrid(heart.list, key = "AnatomicalRegion", legend.position = "right") & scale_color_manual(values = cp) 

```

```{r fig.height=8, fig.width=6}
VlnPlot(
  heart.list.merged,
  group.by = "Sample",
  features = c(
    "nCount_STARsolo",
    "nFeature_STARsolo",
    "nCount_kallisto_collapsed",
    "nFeature_kallisto_collapsed"
  ),
  pt.size = 0,
  ncol = 2
) & theme_classic() & theme(legend.position = "none", text = element_text(size = 12, colour = "black"), axis.text.x = element_text(angle = 45, hjust = 1.0))

```

# Save list of seurat objects, and each object as an adata individually
```{r}
save(
  heart.list,
  file = "/workdir/mm2937/polyAtotal/robjs/heart_list_v4.RData"
)
cat("Done.\n\n")
```

```{r}
save(
  heart.list.merged,
  file = "/workdir/mm2937/polyAtotal/robjs/heart_list_merged_v4.RData"
)
cat("Done.\n\n")
```

```{r}
load("/workdir/mm2937/polyAtotal/robjs/heart_list_v4.RData"
)
load("/workdir/mm2937/polyAtotal/robjs/heart_list_merged_v4.RData"
)
cat("Done.\n\n")
```


```{r}
table(gene_info_gtf[gene_info_gtf$GeneSymbol %in% intersect(rownames(heart.list.merged), gene_info_gtf$GeneSymbol), "biotype_groups"])

biotypes = c("protein_coding")
genelist = intersect(rownames(heart.list.merged), gene_info_gtf$GeneSymbol[!gene_info_gtf$biotype_groups %in% biotypes])
print(length(genelist))

yPAP_genes_mock = FindMarkers(
  heart.list.merged,
  features = genelist,
  group.by = "Sample",
  ident.1 = "yPAP-Mock",
  ident.2 = "Control-Mock",
  logfc.threshold = 0.0,
  only.pos = T
)
yPAP_genes_infected = FindMarkers(
  heart.list.merged,
  features = genelist,
  group.by = "Sample",
  ident.1 = "yPAP-Infected",
  ident.2 = "Control-Infected",
  logfc.threshold = 0.0,
  only.pos = T
)
cat("Done.\n\n")

table(heart.list.merged$polyA)
yPAP_genes_total <- FindMarkers(
  heart.list.merged,
  features = genelist,
  group.by = "polyA",
  ident.1 = "yPAP",
  ident.2 = "ctrl",
  logfc.threshold = 0.0,
  only.pos = T
)

```

```{r}

gene_info_gtf <- gene_info_gtf[!duplicated(gene_info_gtf$GeneSymbol), ]
rownames(gene_info_gtf) <- gene_info_gtf$GeneSymbol
yPAP_genes_total$biotype = gene_info_gtf[rownames(yPAP_genes_total), "biotype_groups"]

yPAP_genes_total = yPAP_genes_total[yPAP_genes_total$avg_log2FC > 0.2,]
yPAP_genes_total = yPAP_genes_total[yPAP_genes_total$p_val_adj < 0.01,]

table(yPAP_genes_total$biotype)
# yPAP_genes_total = yPAP_genes_total[yPAP_genes_total$pct.1 - yPAP_genes_total$pct.2 > 0.3,]
table(yPAP_genes_total$biotype)
yPAP_genes_total[yPAP_genes_total$biotype %in% c("misc_RNA"),]


EnhancedVolcano(
  yPAP_genes_total,
  lab = rownames(yPAP_genes_total),
  x = 'avg_log2FC',
  y = 'p_val_adj',
  title = NULL,
  subtitle = NULL,
  caption = NULL,
  legendPosition = "none",
  pCutoff = 10 ^ -2,
  labSize = 3,
  FCcutoff = 0.0,
  gridlines.major = FALSE,
  gridlines.minor = FALSE,
  colAlpha = 0.8
) + theme_classic() + theme(legend.position = "none") + xlim(0.0, 5.0)
```

```{r}

table(gene_info_gtf$biotype_groups)
biotypes = c("protein_coding", "TCR_genes", "IG_genes")
genelist = gene_info_gtf$GeneSymbol[!gene_info_gtf$biotype_groups %in% biotypes]
length(genelist)

heart.list.filtered <- lapply(
  heart.list, genelist,
  FUN = function(SEU, genelist) 
    return(SEU[genelist,])
)

heart.list.filtered <- lapply(
  heart.list.filtered,
  FUN = function(SEU) seuPreProcess(
    SEU,
    assay="kallisto_collapsed",
    nfeatures = 2000,
    res = 0.6,
    verbose=F
  )
)

heart.list.filtered <- lapply(
  heart.list.filtered,
  FUN = function(SEU) FindClusters(
    SEU,
    assay="kallisto_collapsed",
    res = 1.0,
    verbose=F
  )
)

dim(heart.list[[4]])
dim(heart.list[[2]])

ctrl_genes <- unique(c(rownames(heart.list[[1]]), rownames(heart.list[[2]])))
yPAP_genes <- unique(c(rownames(heart.list[[3]]), rownames(heart.list[[4]])))

genelist = yPAP_genes[!yPAP_genes %in% ctrl_genes]
length(genelist)

heart.list.filtered <- lapply(
  heart.list[c(3,4)], genelist,
  FUN = function(SEU, genelist) 
    return(SEU[genelist,])
)

heart.list.filtered <- lapply(
  heart.list.filtered,
  FUN = function(SEU) seuPreProcess(
    SEU,
    assay="kallisto_collapsed",
    nfeatures = 2000,
    res = 0.6,
    verbose=F
  )
)

heart.list.filtered <- lapply(
  heart.list.filtered,
  FUN = function(SEU) FindClusters(
    SEU,
    assay="kallisto_collapsed",
    res = 1.0,
    verbose=F
  )
)


```


```{r fig.height=5, fig.width=6}


cp <- c("#5AA878", "#E9DF4F", "#3670B7", "#D85B2A", "#C66FA9", "#BDBEC0", "#E3932E", "#60B8EF")
DimPlotGrid(c(heart.list[c(3,4)], heart.list.filtered[c(1,2)]), key = "kallisto_collapsed_snn_res.1", legend.position = "right")

```




## Find genes of given biotypes that are enriched in infected samples
```{r}

table(gene_info_gtf[gene_info_gtf$GeneSymbol %in% intersect(rownames(heart.list.merged), gene_info_gtf$GeneSymbol), "biotype_groups"])

biotypes = c("protein_coding")
genelist = intersect(rownames(heart.list.merged), gene_info_gtf$GeneSymbol[!gene_info_gtf$biotype_groups %in% biotypes])
print(length(genelist))

infection_markers_control = FindMarkers(
  heart.list.merged,
  features = genelist,
  group.by = "Sample",
  ident.1 = "Control-Infected",
  ident.2 = "Control-Mock",
  logfc.threshold = 0.0,
  only.pos = T
)
infection_markers_yPAP = FindMarkers(
  heart.list.merged,
  features = genelist,
  group.by = "Sample",
  ident.1 = "yPAP-Infected",
  ident.2 = "yPAP-Mock",
  logfc.threshold = 0.0,
  only.pos = T
)
cat("Done.\n\n")
```

```{r}

gene_info_gtf <- gene_info_gtf[!duplicated(gene_info_gtf$GeneSymbol), ]
rownames(gene_info_gtf) <- gene_info_gtf$GeneSymbol
infection_markers_yPAP$biotype = gene_info_gtf[rownames(infection_markers_yPAP), "biotype_groups"]

table(infection_markers_yPAP$biotype)

EnhancedVolcano(
  infection_markers_yPAP,
  lab = rownames(infection_markers_yPAP),
  x = 'avg_log2FC',
  y = 'p_val_adj',
  title = NULL,
  subtitle = NULL,
  caption = NULL,
  legendPosition = "none",
  pCutoff = 10 ^ -2,
  labSize = 3,
  FCcutoff = 0.0,
  gridlines.major = FALSE,
  gridlines.minor = FALSE,
  colAlpha = 0.8
) + theme_classic() + theme(legend.position = "none") + xlim(0.0, 1.2)
```

```{r}

infection_markers_yPAP = infection_markers_yPAP[infection_markers_yPAP$avg_log2FC > 0.5,]
infection_markers_yPAP = infection_markers_yPAP[infection_markers_yPAP$p_val_adj < 0.01,]
infection_markers_yPAP = infection_markers_yPAP[infection_markers_yPAP$pct.1 - infection_markers_yPAP$pct.2 > 0.2,]
table(infection_markers_yPAP$biotype)
infection_markers_yPAP[infection_markers_yPAP$biotype %in% c("unprocessed_pseudogene"),]

visListPlot(
  heart.list,
  sample.titles = stringr::str_remove_all(meta_heart$sample,pattern = "Vis_") %>%
    # stringr::str_remove_all(pattern ="yPAP-")%>%
    # stringr::str_remove_all(pattern ="ctrl-")%>%
    stringr::str_remove_all(pattern ="Heart-"),
  assay="kallisto_collapsed",
  reduction="space",
  # slot = 'counts',
  pt.size=0.8,
  legend.position = "bottom",
  font.size = 12,
  axis.title.angle.y=0,
  nrow = 1,
  # combine = F,
  verbose=F,
  features = rownames(infection_markers_yPAP)[infection_markers_yPAP$biotype %in% c("lncRNA")][1:6]
  # reo.genes
) & coord_fixed(ratio = 1.8)

```

## Find genes of given biotypes that are enriched in myocarditic region of the infected samples

```{r}
biotypes = c("protein_coding")
myocarditic_markers_control = FindMarkers(
  heart.list[[2]],
  features = intersect(rownames(heart.list[[2]]), gene_info_gtf$GeneSymbol[!gene_info_gtf$biotype_groups %in% biotypes]),
  group.by = "AnatomicalRegion",
  ident.1 = "Myocarditic region",
  logfc.threshold = 0.0,
  only.pos = T
)
myocarditic_markers_yPAP = FindMarkers(
  heart.list[[4]],
  features = intersect(rownames(heart.list[[4]]), gene_info_gtf$GeneSymbol[!gene_info_gtf$biotype_groups %in% biotypes]),
  group.by = "AnatomicalRegion",
  ident.1 = "Myocarditic region",
  logfc.threshold = 0.0,
  only.pos = T
)
cat("Done.\n\n")

```




```{r}

gene_info_gtf <- gene_info_gtf[!duplicated(gene_info_gtf$GeneSymbol), ]
rownames(gene_info_gtf) <- gene_info_gtf$GeneSymbol
myocarditic_markers_yPAP$biotype = gene_info_gtf[rownames(myocarditic_markers_yPAP), "biotype_groups"]

myocarditic_markers_yPAP = myocarditic_markers_yPAP[myocarditic_markers_yPAP$avg_log2FC > 0.2,]
myocarditic_markers_yPAP = myocarditic_markers_yPAP[myocarditic_markers_yPAP$p_val_adj < 0.01,]
# myocarditic_markers_yPAP = myocarditic_markers_yPAP[myocarditic_markers_yPAP$pct.1 - myocarditic_markers_yPAP$pct.2 > 0.2,]
table(myocarditic_markers_yPAP$biotype)
myocarditic_markers_yPAP[myocarditic_markers_yPAP$biotype %in% c("polymorphic_pseudogene"),]

table(myocarditic_markers_yPAP$biotype)

EnhancedVolcano(
  myocarditic_markers_yPAP,
  lab = rownames(myocarditic_markers_yPAP),
  x = 'avg_log2FC',
  y = 'p_val_adj',
  title = NULL,
  subtitle = NULL,
  caption = NULL,
  legendPosition = "none",
  pCutoff = 10 ^ -2,
  labSize = 3,
  FCcutoff = 0.0,
  gridlines.major = FALSE,
  gridlines.minor = FALSE,
  colAlpha = 0.8
) + theme_classic() + theme(legend.position = "none") + xlim(0.0, 1.5)
```



```{r}
DefaultAssay(heart.list[[4]])
visListPlot(
  heart.list[c(4)],
  sample.titles = stringr::str_remove_all(meta_heart$sample,pattern = "Vis_") %>%
    # stringr::str_remove_all(pattern ="yPAP-")%>%
    # stringr::str_remove_all(pattern ="ctrl-")%>%
    stringr::str_remove_all(pattern ="Heart-"),
  assay="kallisto_collapsed",
  reduction="space",
  # slot = 'counts',
  pt.size=0.8,
  legend.position = "bottom",
  font.size = 12,
  axis.title.angle.y=0,
  nrow = 1,
  # combine = F,
  verbose=F,
  features = rownames(myocarditic_markers_yPAP)[myocarditic_markers_yPAP$biotype %in% c("polymorphic_pseudogene")]
  # reo.genes
) & coord_fixed(ratio = 1.8)
```



```{r fig.height=4, fig.width=8}

cp <- c("#5AA878", "#E9DF4F", "#3670B7", "#D85B2A", "#C66FA9", "#BDBEC0", "#E3932E", "#60B8EF")

VlnPlot(heart.list.merged[,heart.list.merged$Sample == "yPAP-Infected"], features = rownames(myocarditic_markers_yPAP)[myocarditic_markers_yPAP$biotype %in% c("polymorphic_pseudogene")], group.by = "AnatomicalRegion", pt.size = 0.0, ncol = 2) & theme_classic() & theme(legend.position = "none", axis.text.x = element_blank()) & scale_fill_manual(values = cp) 

temp <- heart.list.merged[,heart.list.merged$Sample == "yPAP-Infected"]
VlnPlot(temp, features = c("nCount_xGen.kallisto.sense.log2p1", "nCount_xGen.kallisto.as.log2p1"), group.by = "AnatomicalRegion", pt.size = 0.1, ncol = 2) & theme_classic() & theme(legend.position = "none", axis.text.x = element_blank()) & scale_fill_manual(values = cp) 


```


```{r}
library(monocle)

infected_yPAP_CDS <- as.CellDataSet(heart.list[[4]])
infected_yPAP_CDS@expressionFamily <- negbinomial.size()
# pData(infected_yPAP_CDS)
# fData(infected_yPAP_CDS)

non_cavity_cells <- 
  rownames(pData(infected_yPAP_CDS))[pData(infected_yPAP_CDS)["AnatomicalRegion"] != "Cavity"]
infected_yPAP_CDS <- infected_yPAP_CDS[, non_cavity_cells]
dim(infected_yPAP_CDS)

```

## Createing and processing CDS
```{r}

infected_yPAP_CDS <- estimateSizeFactors(infected_yPAP_CDS)
infected_yPAP_CDS <- estimateDispersions(infected_yPAP_CDS)

infected_yPAP_CDS <- detectGenes(infected_yPAP_CDS, min_expr = 0.001)
dim(infected_yPAP_CDS)

expressed_genes <- row.names(subset(fData(infected_yPAP_CDS),
    num_cells_expressed >= 10))
infected_yPAP_CDS <- infected_yPAP_CDS[expressed_genes,]
dim(infected_yPAP_CDS)

```

## Selecting genes for ReovirusDE
```{r}
disp_table <- dispersionTable(infected_yPAP_CDS)

# ordering_genes <- subset(disp_table,
#                          mean_expression >= 0.01 &
#                            dispersion_empirical >= 0.5 * dispersion_fit)$gene_id
# length(ordering_genes)
# infected_yPAP_CDS <-
#   setOrderingFilter(infected_yPAP_CDS, ordering_genes = ordering_genes)
# plot_ordering_genes(infected_yPAP_CDS)

table(gene_info_gtf$biotype_groups)
biotypes = c("protein_coding", "IG_genes", "TCR_genes")
ordering_genes <- intersect(disp_table$gene_id, gene_info_gtf$GeneSymbol[!gene_info_gtf$biotype_groups %in% biotypes])
length(ordering_genes)
infected_yPAP_CDS <-
  setOrderingFilter(infected_yPAP_CDS, ordering_genes = ordering_genes)

```

# Finding genes that change with reovirus counts
```{r}

dim(infected_yPAP_CDS)
CDS_subset <-
  infected_yPAP_CDS[ordering_genes[ordering_genes %in% rownames(infected_yPAP_CDS)], ]
dim(CDS_subset)

pData(CDS_subset)[, "Pseudotime"] <-
  pData(CDS_subset)$nCount_xGen.kallisto.log2p1
diff_test_res <- differentialGeneTest(CDS_subset,
                                      fullModelFormulaStr = "~sm.ns(Pseudotime)", cores = 20)

write.csv(diff_test_res, file = "./diff_test_res.non.protein.coding.csv")

dim(diff_test_res)
sig_gene_names <- subset(diff_test_res, qval < 0.01)
sig_gene_names <- rownames(sig_gene_names[order(sig_gene_names$qval),])
print(sig_gene_names)




```

## Plot GE trends
```{r}

((FeatureScatter(heart.list[[4]], feature1 = "nCount_xGen.kallisto.log2p1", feature2 = sig_gene_names[1], span = T) | FeatureScatter(heart.list[[4]], feature1 = "nCount_xGen.kallisto.log2p1", feature2 = sig_gene_names[2], span = T)) / (FeatureScatter(heart.list[[4]], feature1 = "nCount_xGen.kallisto.log2p1", feature2 = sig_gene_names[3], span = T) | FeatureScatter(heart.list[[4]], feature1 = "nCount_xGen.kallisto.log2p1", feature2 = sig_gene_names[4], span = T))) & theme(legend.position = "none") & ggtitle(NULL) & scale_color_manual(values = cp)

```


```{r fig.height=5, fig.width=6}
cov_plus <- as.data.frame(t(read.csv("./../data/coverage_plts/reovirus_plus.geneBodyCoverage.txt", sep = "\t", row.names = 1)))
rownames(cov_plus) <- str_split_fixed(rownames(cov_plus), "X", 2)[,2]
cov_plus <- apply(cov_plus,2,function(x){x/max(x)})
cov_plus$Percentile <- rownames(cov_plus)             
cov_plus

cov_minus <- as.data.frame(t(read.csv("./../data/coverage_plts/reovirus_minus.geneBodyCoverage.txt", sep = "\t", row.names = 1)))
cov_minus <- as.data.frame(apply(cov_minus,2,function(x){x/max(x, na.rm = T)}))
rownames(cov_minus) <- rev(str_split_fixed(rownames(cov_minus), "X", 2)[,2])
cov_minus$Percentile <- rownames(cov_minus)             

cov <- as.data.frame(merge(cov_plus, cov_minus, by = "Percentile"))
cov$Percentile <- as.numeric(cov$Percentile)
cov = cov[order(cov$Percentile),]

library(reshape2)
cov_data <- melt(cov, id.vars = "Percentile", variable.name = "Sample", value.name = "Coverage")
cov_data$Strand <- str_split_fixed(cov_data$Sample, pattern = "_xGen.", 2)[,2]
cov_data$Sample <- str_split_fixed(cov_data$Sample, pattern = "_xGen.", 2)[,1]
cov_data$Coverage[cov_data$Strand == "minus"] = -1 * cov_data$Coverage[cov_data$Strand == "minus"]

((ggplot(cov_data[cov_data$Strand == "plus",], mapping = aes(x = Percentile, y = Coverage, color = Sample)) + geom_line() + ylim(0, max(cov_data$Coverage))) / ggplot(cov_data[cov_data$Strand == "minus",], mapping = aes(x = Percentile, y = Coverage, color = Sample)) + geom_line() + scale_y_reverse() + ylim(-1 * max(cov_data$Coverage), 0)) & xlab("Percentile 5' -> 3'") & theme_classic()

```


# Session Info
```{r}
sessionInfo()
```
