#  Comparison of standard visium to polyA visium in notexin-injured tibialis anterior muscles
## David W. McKellar

# Session setup
## Libs, setwd
```{r message=FALSE, warning=FALSE}
library(Matrix)
library(dplyr)
library(Seurat)
library(future)

library(ggplot2)
library(patchwork)
library(viridis)
library(data.table)
library(shades)

source("/home/dwm269/DWM_utils/plotting_utils/scThemes.R")
source("/home/dwm269/DWM_utils/sc_utils/seurat_helpers/seutils.R")
source("/home/dwm269/DWM_utils/sc_utils/seurat_helpers/seuplots.R")
```

## Figure settings & colors
```{r}
# fonts, sizes, etc.
small.font = 6*2
big.font = 8*2
line.width = 0.5
pt.size=0.01
pt.stroke=0.3
label.size=2
```

## Plot themes 
```{r}
scTheme <- scThemes(
  small.font = small.font,
  big.font = big.font,
  line.width = line.width,
  pt.size=pt.size,
  pt.stroke=pt.stroke,
  label.size=label.size
)

print(names(scTheme))
```
```{r}
mckolors <- read.csv("/home/dwm269/DWM_utils/plotting_utils/McKolors_v1.csv") %>% 
  as.list() %>%
  lapply(
    FUN=function(X) X[X!=""]
  )
names(mckolors)
```

# Load and subset preprocessed Visium +/- spTotal data
```{r}
load("/local/workdir/dwm269/totalRNA/spTotal/robjs/vis_list_v3.RData")
cat("Done.")
```

# Load metadata
```{r}
meta_vis <- read.csv("/workdir/dwm269/totalRNA/spTotal/resources/meta_sheet.csv")
meta_skm <- meta_vis[meta_vis$tissue=="muscle"&meta_vis$rnase_inhib%in%c("ctrl", "Protector"),] #subset metadata

skm.list <- vis.list[meta_vis$tissue=="muscle"&meta_vis$rnase_inhib%in%c("ctrl", "Protector")] # subset out Seurat objects
skm.list <- skm.list[with(meta_skm, order(meta_skm$polyA,meta_skm$timepoint))] # reorder list for visualization

meta_skm <- meta_skm[with(meta_skm, order(meta_skm$polyA,meta_skm$timepoint)),] #sort metadata (by preprocessing, then by injury timepoint)

meta_skm
```

# STARsolo vs. kallisto/BUStools
```{r warning=FALSE}
suppressMessages(
  visListPlot(
    skm.list,
    sample.titles = stringr::str_remove_all(meta_skm$sample,pattern = "Vis_") %>%
      stringr::str_remove_all(pattern ="_SkM"),
    reduction="space",
    pt.size=0.4,
    legend.position = "bottom",
    font.size = 6,
    axis.title.angle.y=0,
    nrow = 1,
    combine = T,
    verbose=F,
    colormap = mckolors$Spectral %>% rev(),
    features = c(
      "nFeature_STARsolo_collapsed","nFeature_kallisto_collapsed",
      "nCount_STARsolo_collapsed","nCount_kallisto_collapsed"
    )
  )&theme(
    legend.text = element_text(size=6)
  )&coord_fixed(ratio=1/1.6)
)
```

Scatter plot for each spot, colored by preproceessing, to look at # Features and # UMIs detected by STARsolo vs. kallisto
```{r fig.height=6, fig.width=12}
if(!exists("skm.merged")){
  skm.merged <- merge(
    skm.list[[1]],
    skm.list[2:length(skm.list)],
    add.cell.ids = meta_skm$sample
  )
}
wrap_plots(
  ggplot(
    skm.merged@meta.data[sample(rownames(skm.merged@meta.data)),],
    aes(
      x=nFeature_STARsolo,
      y=nFeature_kallisto,
      color=polyA
    )
  )+
    geom_point(alpha=0.5)+
    geom_abline()+
    scTheme$fig1bcd+
    scale_color_manual(values=mckolors$txg)+
    guides(color=guide_legend(override.aes = list(size=4,alpha=1,fill=NA))),
  
  ggplot(
    skm.merged@meta.data[sample(rownames(skm.merged@meta.data)),],
    aes(
      x=nCount_STARsolo,
      y=nCount_kallisto,
      color=polyA
    )
  )+
    geom_point(alpha=0.5)+
    geom_abline()+
    scTheme$fig1bcd+
    scale_color_manual(values=mckolors$txg)+
    guides(color=guide_legend(override.aes = list(size=4,alpha=1,fill=NA))),
  guides = "collect"
)&scTheme$scatter&theme(
  legend.position="right"
)
```
# Knee plots by assay
```{r}
lapply(
  skm.list,
  FUN=function(VIS){
    qplot(
      1:length(VIS$nCount_kallisto),
      sort(VIS$nCount_kallisto, decreasing=T)
    )+
      scTheme$scatter+
      theme(axis.title.x = element_blank())+
      labs(
        y="nCount_kallisto",
        title=VIS$sample[1]
      )+
      scale_y_continuous(limits = c(0,75000))
  }
)%>%
  wrap_plots()&scTheme$scatter

```

# Get gene biotype info from the GENCODE annotations
`gtf.info` is just a reformatted .tsv of the GENCODE M28 annotations .gtf
```{r}
gtf.info <- read.csv(
  file="/workdir/dwm269/totalRNA/spTotal/resources/gene_lists/GRCm39_GENCODEm28_gene_info_gtf.tsv",
  sep = "\t"
)

# Remove period/decimal suffixes from ensembl IDs
gtf.info$GeneID <- strsplit(
  x=gtf.info$GENEID,
  split = "\\."
)%>%
  lapply(
    FUN=function(X) X[[1]]
  )%>%
  unlist()

cat(
  paste("Using", length(unique(gtf.info$GeneSymbol)), "genes, and ",length(unique(gtf.info$Biotype)),"biotypes...\n")
)

head(gtf.info)
```

## gene lists by biotype
```{r}
genes = list()
for(biotype in unique(gtf.info$Biotype)){
  genes[[biotype]] <- gtf.info$GENEID[gtf.info$Biotype==biotype] %>%
    ens2gene(
      biomart.info = gtf.info,
      ens.colname = "GENEID",
      gene.colname = "GeneSymbol"
      ) %>%
    unique()
}

names(genes)
```

## Gene biotype spatial plots
```{r warning=FALSE}
visListPlot(
  skm.list,
  sample.titles = stringr::str_remove_all(meta_skm$sample,pattern = "Vis_") %>%
        stringr::str_remove_all(pattern ="yPAP_")%>%
        stringr::str_remove_all(pattern ="ctrl_")%>%
        stringr::str_remove_all(pattern ="_SkM"),
  # assay="STARsolo_collapsed",
  reduction="space",
  # slot = 'counts',
  pt.size=0.6,
  legend.position = "bottom",
  font.size = 12,
  axis.title.angle.y=0,
  nrow = 1,
  combine = T,
  verbose=F,
  # colormap="inferno",colormap.direction = -1,
  colormap = mckolors$RdYlBu%>%rev(),
  features = c(
    # "nFeature_STARsolo","nCount_STARsolo","percent.mt",
    "pct.protein_coding",
    # "pct.lncRNA",
    "pct.rRNA",
    # "pct.Mt_rRNA", 
    "pct.Mt_tRNA",
    # "pct.ribo.pp",
    "pct.snoRNA",
    "pct.miRNA",
    "pct.snRNA"
  ) 
)%>% suppressMessages()
```

# Print highly expressed miRNAs for each object
```{r}
trnas <- mouse.info$ensembl_gene_id[mouse.info$gene_biotype=="miRNA"]
top.trnas <- lapply(
  skm.list,
  FUN = function(VIS) grepGenes(
    VIS,
    assay="kallisto",
    pattern = mirnas,
    verbose = F
  ) 
)

lapply(
  top.trnas,
  FUN = function(X) X %>%
    head(n=20) %>%
    print()
)
```
# Print highly expressed tRNAs for each object
```{r}
top.mirs <- lapply(
  skm.list,
  FUN = function(VIS) grepGenes(
    VIS,
    assay="kallisto_collapsed",
    pattern = genes$miR,
    filter.pattern = c("hg","ENS")
  ) 
)

lapply(
  top.mirs,
  FUN = function(X) X %>%
    head(n=20) %>%
    print()
)
```

# Gene spatial feature plots
## Canonical myogenesis markers
```{r fig.height=10, fig.width=10}
suppressMessages(
  visListPlot(
    skm.list,
    sample.titles = stringr::str_remove_all(meta_skm$sample,pattern = "Vis_") %>%
      stringr::str_remove_all(pattern ="_SkM"),
    reduction="space",
    assay="kallisto_collapsed", slot = 'data',
    pt.size=0.5,
    legend.position = "bottom",
    font.size = 6,
    axis.title.angle.y=0,
    nrow = 1,
    combine = T,
    verbose=F,
    colormap = "plasma",
    # colormap = mckolors$Spectral%>%rev(),
    features = c(
      "Myf5","Myod1","Myog","Myh1","Myh4"
    )
  )&theme(
    legend.text = element_text(size=6)
  )&coord_fixed(
    ratio=1/1.6
  )
)
```

```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
suppressMessages(
  visListPlot(
    skm.list,
    sample.titles = stringr::str_remove_all(meta_skm$sample,pattern = "Vis_") %>%
      stringr::str_remove_all(pattern ="_SkM"),
    reduction="space",
    assay="kallisto_collapsed", 
    slot = 'counts',
    pt.size=0.5,
    legend.position = "bottom",
    font.size = 6,
    axis.title.angle.y=0,
    nrow = 1,
    combine = T,
    verbose=F,
    colormap = "plasma",
    # colormap = mckolors$Spectral%>%rev(),
    features = c(
      "Mir703","Mir133a-2","Mir1b","Mir1291","Mir351"
    )
  )&theme(
    legend.text = element_text(size=6)
  )&coord_fixed(
    ratio=1/1.6
  )
)
```


# Clustering analysis
Multi-resolution clustering 
```{r}
skm.list <- skm.list %>% 
  lapply(
    FUN=function(VIS){
      for(RES in c(0.2,0.4,0.6,0.8)){
        VIS = FindClusters(
          object=VIS,
          graph="kallisto_collapsed_snn",
          resolution=RES,
          verbose=T
        )
      }
      return(VIS)
    }
  )
```

```{r}
lapply(
  skm.list,
  FUN=function(VIS) DimPlot(
    VIS,
    reduction = "space",
    group.by = c(
      "kallisto_collapsed_snn_res.0.4"
    )
  )
) %>% 
  lapply(
    FUN=function(VIS) VIS +
      scTheme$umap +
      theme(
        axis.title = element_blank(),
        axis.line =  element_blank(),
        plot.title=element_blank()
      )
  ) %>% wrap_plots()
```

# Spatially variable features
```{r}
library(parallel)
num.cores=9
cl <- makeCluster(num.cores)

skm.list <- mclapply(
  skm.list,
  FUN = function(SEU) FindSpatiallyVariableFeatures(
      SEU,
      assay = "kallisto_collapsed", 
      # features = VariableFeatures(SEU)[1:1000],
      selection.method = "markvariogram"
  ),
  mc.cores=num.cores
)

```


```{r}
top.spatial <- lapply(
  skm.list,
  FUN=function(VIS) SpatiallyVariableFeatures(
    VIS, selection.method = "markvariogram"
  ) %>%
    head(n=50)
) 

tmp.spatial <- top.spatial %>%
  unlist() %>%
  unique()
# tmp.spatial = tmp.spatial[tmp.spati]
tmp.spatial
```


```{r fig.height=6, fig.width=16, message=FALSE, warning=FALSE}
suppressMessages(
  visListPlot(
    skm.list,
    sample.titles = stringr::str_remove_all(meta_skm$sample, pattern = "Vis_") %>%
      stringr::str_remove_all(pattern ="_SkM"),
    reduction="space",
    assay="kallisto_collapsed",
    slot = 'data',
    pt.size=0.25,
    legend.position = "bottom",
    font.size = 8,
    axis.title.angle.y=0,
    colormap = mckolors$cubs,
    nrow = 1,
    combine = T,
    verbose=F,
    
    # features = tmp.spatial[tmp.spatial %in% genes.ncrna] %>%rev()%>% head(n=12)
    # features=grepGenes(skm.list[[7]],assay="kallisto_collapsed",pattern = "Mir",filter.pattern = "hg")[1:12]
    features = grepGenes(skm.list[[6]],assay="kallisto_collapsed",pattern = genes$miRNA,filter.pattern = "ENS")[1:12]
    # features=grepGenes(skm.list[[6]],assay="kallisto_collapsed",pattern = c("Sno","SNO"))[1:12]
  )&theme(
    legend.text = element_text(size=6)
  )
)
```

# polyA differentially captured features
```{r}


```

# Spliced vs. unspliced analysis
```{r}
require("R.utils")
NCORES=12
for(i in 1:length(skm.list)){
  
  # Spliced ----
  tmp.seu <- ReadMtx(
    mtx=paste0(meta_skm$data.dir.kallisto[i],"/kb_lamanno/counts_unfiltered/spliced.mtx"),
    cells=paste0(meta_skm$data.dir.kallisto[i],"/kb_lamanno/counts_unfiltered/spliced.barcodes.txt"),
    features=paste0(meta_skm$data.dir.kallisto[i],"/kb_lamanno/counts_unfiltered/spliced.genes.txt"),
    feature.column=1,
    mtx.transpose = T
  ) %>%
    CreateSeuratObject() %>%
    subset( # select cells
      cells=Cells(skm.list[[i]])
    )
  
  # Select non-zero features
  tmp.seu <- subset( 
    tmp.seu,
    features = rownames(tmp.seu)[Matrix::rowSums(tmp.seu) > 0]
  )
  
  # convert ensembl IDs to MGI gene IDs
  tmp.mat <- GetAssayData(tmp.seu, assay="RNA")
  rownames(tmp.mat) <- ens2gene(
    ens=rownames(tmp.mat),
    biomart.info = gtf.info,
    ens.colname = "GENEID", # Ensembl ID, yes periods
    # ens.colname = "GeneID", # ensembl IDs, no periods
    gene.colname = "GeneSymbol",
    force.unique = T,
    ncores=NCORES,
    verbose=T
  )
  
  # add kallisto counts to vis_list
  skm.list[[i]][["spliced"]] <- CreateAssayObject(
    counts = tmp.mat,
    min.cells = 1
  )
  
  
  # Unspliced ----
  tmp.seu <- ReadMtx(
    mtx=paste0(meta_skm$data.dir.kallisto[i],"/kb_lamanno/counts_unfiltered/unspliced.mtx"),
    cells=paste0(meta_skm$data.dir.kallisto[i],"/kb_lamanno/counts_unfiltered/unspliced.barcodes.txt"),
    features=paste0(meta_skm$data.dir.kallisto[i],"/kb_lamanno/counts_unfiltered/unspliced.genes.txt"),
    feature.column=1,
    mtx.transpose = T
  ) %>%
    CreateSeuratObject() %>%
    subset( # select cells
      cells=Cells(skm.list[[i]])
    )
  
  # Select non-zero features
  tmp.seu <- subset( 
    tmp.seu,
    features = rownames(tmp.seu)[Matrix::rowSums(tmp.seu) > 0]
  )
  
  # convert ensembl IDs to MGI gene IDs
  tmp.mat <- GetAssayData(tmp.seu, assay="RNA")
  rownames(tmp.mat) <- ens2gene(
    ens=rownames(tmp.mat),
    biomart.info = gtf.info,
    ens.colname = "GENEID", # Ensembl ID, yes periods
    # ens.colname = "GeneID", # ensembl IDs, no periods
    gene.colname = "GeneSymbol",
    force.unique = T,
    ncores=NCORES,
    verbose=T
  )
  
  # add kallisto counts to vis_list
  skm.list[[i]][["unspliced"]] <- CreateAssayObject(
    counts = tmp.mat,
    min.cells = 1
  )
}

rm(tmp.mat)
rm(tmp.seu)
gc()
```
Normalize the spliced/unspliced read counts for visualization
```{r}
skm.list <- lapply(
  skm.list,
  FUN=function(SEU){
    SEU <- NormalizeData(
      SEU,
      assay="spliced"
    )
    
    SEU <- NormalizeData(
      SEU,
      assay="unspliced"
    )
    
    return(SEU)
  }
)
message("Done.")
```

Visualize gene splicing ratios with visCoMap
```{r fig.height=8, fig.width=6}
tmp.feat = "Myf6"
visCoMap(
    skm.list,
    sample.titles = meta_skm$sample,
    reduction="space",
    slot = c('data','data'),
    pt.size=0.6,
    legend.position = "bottom",
    font.size = 8,
    axis.title.angle.y=0,
    nrow = 1,
    combine = T,
    verbose=T,
    # colormap = mckolors$Spectral,
    # colormap.direction = 1,
    comap.fxn = function(X) X[1]/X[2],
    coex.name = "Unspliced/Total", alt.titles = paste(tmp.feat,c("Unspliced", "Total")),
    features = c(tmp.feat, tmp.feat),assay=c("unspliced","kallisto_collapsed"),
  )&coord_fixed(ratio=1/1.6)
```

# Session Info
```{r}
sessionInfo()
```