#  Comparison of standard visium to polyA visium in notexin-injured tibialis anterior muscles
## David W. McKellar

# Session setup
## Libs, setwd
```{r message=FALSE, warning=FALSE}
library(Matrix)
library(dplyr)
library(Seurat)
library(future)

library(ggplot2)
library(patchwork)
library(viridis)
library(data.table)
library(shades)

source("/home/dwm269/DWM_utils/plotting_utils/scThemes.R")
source("/home/dwm269/DWM_utils/sc_utils/seurat_helpers/seutils.R")
source("/home/dwm269/DWM_utils/sc_utils/seurat_helpers/seuplots.R")
```

## Figure settings & colors
```{r}
# fonts, sizes, etc.
small.font = 6*2
big.font = 8*2
line.width = 0.5
pt.size=0.01
pt.stroke=0.3
label.size=2
```

## Plot themes 
```{r}
scTheme <- scThemes(
  small.font = small.font,
  big.font = big.font,
  line.width = line.width,
  pt.size=pt.size,
  pt.stroke=pt.stroke,
  label.size=label.size
)

print(names(scTheme))
```

```{r}
mckolors <- read.csv("/home/dwm269/DWM_utils/plotting_utils/McKolors_v1.csv") %>% 
  as.list() %>%
  lapply(
    FUN=function(X) X[X!=""]
  )
mckolors
```

# Load and subset preprocessed Visium +/- spTotal data
```{r}
load("/local/workdir/dwm269/totalRNA/spTotal/robjs/vis_list_v2.RData")
cat("Done.")
```

# Load metadata
```{r}
meta_vis <- read.csv("/workdir/dwm269/totalRNA/spTotal/resources/meta_sheet.csv")
meta_skm <- meta_vis[meta_vis$tissue=="muscle"&meta_vis$rnase_inhib%in%c("ctrl", "Protector"),]

skm.list <- vis.list[meta_vis$tissue=="muscle"&meta_vis$rnase_inhib%in%c("ctrl", "Protector")]
meta_skm
```

# STARsolo vs. kallisto/BUStools
```{r}
suppressMessages(
  visListPlot(
    skm.list,
    sample.titles = stringr::str_remove_all(meta_skm$sample,pattern = "Vis_") %>%
      stringr::str_remove_all(pattern ="_SkM"),
    reduction="space",
    # slot = 'counts',
    pt.size=0.4,
    legend.position = "bottom",
    font.size = 6,
    axis.title.angle.y=0,
    nrow = 1,
    combine = T,
    verbose=F,
    colormap = mckolors$Spectral %>% rev(),
    features = c(
      "nFeature_STARsolo","nFeature_kallisto",
      "nCount_STARsolo","nCount_kallisto"
    )
  )&theme(
    legend.text = element_text(size=6)
  )
)
```
Scatter plot for each spot, colored by preproceessing, to look at # Features and # UMIs detected by STARsolo vs. kallisto
```{r fig.height=6, fig.width=12}
if(!exists("skm.merged")){
  skm.merged <- merge(
    skm.list[[1]],
    skm.list[2:length(skm.list)],
    add.cell.ids = meta_skm$sample
  )
}
wrap_plots(
  ggplot(
    skm.merged@meta.data[sample(rownames(skm.merged@meta.data)),],
    aes(
      x=nFeature_STARsolo,
      y=nFeature_kallisto,
      color=polyA
    )
  )+
    geom_point(alpha=0.5)+
    geom_abline()+
    scTheme$fig1bcd+
    scale_color_manual(values=mckolors$capitals)+
    guides(color=guide_legend(override.aes = list(size=4,alpha=1,fill=NA))),
  
  ggplot(
    skm.merged@meta.data[sample(rownames(skm.merged@meta.data)),],
    aes(
      x=nCount_STARsolo,
      y=nCount_kallisto,
      color=polyA
    )
  )+
    geom_point(alpha=0.5)+
    geom_abline()+
    scTheme$fig1bcd+
    scale_color_manual(values=mckolors$capitals)+
    guides(color=guide_legend(override.aes = list(size=4,alpha=1,fill=NA))),
  guides = "collect"
)&theme(
  legend.position="right"
)
```
# biomaRt
```{r}
if(!exists("mouse.info")){
  mouse = biomaRt::useMart(
    # biomart = "ENSEMBL_MART_MOUSE",
   biomart = "ENSEMBL_MART_ENSEMBL",
    dataset = "mmusculus_gene_ensembl"
    # host = "https://apr2020.archive.ensembl.org" ##for mm10 info
  )
  mouse.info <- biomaRt::getBM(
    attributes = c(
      "mgi_symbol", "gene_biotype","transcript_biotype", 
      # "description", "name_1006","definition_1006","external_gene_name",
      "ensembl_gene_id","ensembl_transcript_id","ensembl_peptide_id",
      "chromosome_name","transcript_length", "strand"
    ),
    uniqueRows = T,
    # host = "useast.ensembl.org",
    verbose = T,
    mart = mouse
  )
}

cat(
  paste("Using", length(unique(mouse.info$mgi_symbol)), "genes, and ",length(unique(mouse.info$gene_biotype)),"biotypes...\n")
)
```

# Print highly expressed miRNAs for each object
```{r}
trnas <- mouse.info$ensembl_gene_id[mouse.info$gene_biotype=="miRNA"]
top.trnas <- lapply(
  skm.list,
  FUN = function(VIS) grepGenes(
    VIS,
    assay="kallisto",
    pattern = mirnas,
    verbose = F
  ) 
)

lapply(
  top.trnas,
  FUN = function(X) X %>%
    head(n=20) %>%
    print()
)
```
# Print highly expressed tRNAs for each object
```{r}
mirnas <- mouse.info$ensembl_gene_id[mouse.info$gene_biotype %in% c("Mt_tRNA")]
top.mirs <- lapply(
  skm.list,
  FUN = function(VIS) grepGenes(
    VIS,
    assay="kallisto",
    pattern = mirnas
  ) 
)

lapply(
  top.mirs,
  FUN = function(X) X %>%
    head(n=20) %>%
    print()
)
```


# Gene spatial feature plots
```{r}
i=c(6,7,4,5)
suppressMessages(
  visListPlot(
    skm.list[i],
    sample.titles = stringr::str_remove_all(meta_skm$sample[i], pattern = "Vis_") %>%
      stringr::str_remove_all(pattern ="_SkM"),
    reduction="space",
    assay="kallisto",
    slot = 'data',
    pt.size=1,
    legend.position = "bottom",
    font.size = 6,
    axis.title.angle.y=0,
    nrow = 1,
    combine = T,
    verbose=F,
    # colormap = mckolors$Spectral %>% rev(),
    # features = c(
    #   "ENSMUSG00000106348", #Mir1a-2
    #   "ENSMUSG00000065453", #Mirlet7d
    #   "ENSMUSG00000065405", # Mir30a
    #   "ENSMUSG00000065516", # Mir214
    #   "ENSMUSG00000065480" # Mir133b
    # ),
    # alt.titles = c(
    #   "Mir1a-2", #ENSMUSG00000106348
    #   "Mirlet7d", #ENSMUSG00000065453
    #   "Mir30a", #ENSMUSG00000065405
    #   "Mir214", #ENSMUSG00000065516
    #   "Mir133b" #ENSMUSG00000065480
    # )
    
    features = lapply(
      top.trnas[4:7],# top tRNAs in polyA samples...
      function(X) head(X, n=8)
    ) %>% unlist() %>% unique() %>% head(n=10)
  )&theme(
    legend.text = element_text(size=6)
  )
)
```
















# Session Info
```{r}
sessionInfo()
```